### 2017-03-16 abstract.tex
Splitting main to sections

\begin{abstract}
Learning the dependency structure of a multivariate distribution from
the observational data is an important task since it allows us to
speculate about the underlying causal mechanisms that induce
dependencies. This learning task can be approached as a model
selection problem, but recent studies have revealed that the popular
Bayesian model selection criterion is not satisfactory. We review some
of the information theoretic alternatives and introduce a new one
called a quotient normalized maximum likelihood criterion.
\end{abstract}