### 2017-03-16 conclusion.tex

\section{CONCLUSION}

We have presented qNML, a new model selection criterion for learning
structures of Bayesian networks.  While being competitive in
predictive terms, it often yields significantly simpler models than
other common model selection criteria.  The computational cost of qNML
equals the cost of earlier criteria.  The criterion also gives equal
scores for models that encode same independence hypotheses about the
joint probability distribution.  qNML also coincides with NML
criterion for exponentially many models.