### 2017-03-31 abstract.tex
Typos

\begin{abstract}
Learning the dependency structure of a multivariate distribution from
observational data is an important task since it allows us to
speculate about the underlying causal mechanisms that induce
dependencies. This learning task can be approached as a model
selection problem, but recent studies have revealed that the popular
Bayesian model selection criteria are not satisfactory. We review some
of the information-theoretic alternatives and introduce a new criterion
called the quotient normalized maximum likelihood criterion.
\end{abstract}