### 2017-10-12 supp_mat_C.tex

\section{Prediction and Parsimony}

We took 20 UCI data sets created 1000 permutations of each data set.
We then took the first x\% of each permuted data set data (x in
${10,20,30,\ldots,90}$ as training data and used the remaining part as
a test data. We used exact structure learning to learn the best
scoring model using four model selection criteria, BDeu ($\alpha=1$),
BIC, fNML and qNML, recording the number of parameters in the learnt
models and the average predictive log-probability of each test
vector. These numbers are then collected for 1000 different data
permutations average and variance of which appear on the following
pages, one page per data set.