### 2017-10-13 conclusion.tex
Typos

\section{CONCLUSION}

We have presented qNML, a new model selection criterion for learning
structures of Bayesian networks.  While being competitive in
predictive terms, it often yields significantly simpler models than
other common model selection criteria other than BIC that has a very strong
bias for simplicity. The computational cost of qNML
equals the cost of the current state-of-the-art criteria. The
criterion also gives equal scores for models that encode the same
independence hypotheses about the joint probability distribution.
qNML also coincides with the NML criterion for many models.
In our experiments, the qNML criterion appears as a safe choice for a
model selection criterion that balances
parsimony, predictive capability and the ability to quickly converge to the
generating model.