### 2018-02-14 main.tex
[accepted] options + authot info

%\documentclass[letterpaper]{article}
%\usepackage{proceed2e}
%\usepackage[margin=1in]{geometry}

\documentclass[twoside]{article}
\usepackage[accepted]{aistats2018}
\usepackage{amsmath,amsthm,epsfig,booktabs,multirow}
\usepackage[hidelinks]{hyperref}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newcommand{\Le}{\left(}
\newcommand{\Ri}{\right)}

% Set the typeface to Times Roman
%\usepackage{times}

%\title{Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures}

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

\hyphenation{col-umns}
\hyphenation{Bayes-ian}
\hyphenation{oth-er}

\runningauthor{Tomi Silander, Janne Lepp\"{a}-aho, Elias J\"{a}\"{a}saari, Teemu Roos}

\begin{document}

\twocolumn[

\aistatstitle{Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures}

\aistatsauthor{ Tomi Silander$^1$ \And Janne Lepp\"{a}-aho$^{2,4}$ \And Elias J\"{a}\"{a}saari$^{2,3}$ \And Teemu Roos$^{2,4}$   }

\aistatsaddress{ $^1$NAVER LABS Europe \And $^2$Helsinki Institute for \\ Information Technology HIIT \And $^3$Department of CS,\\Aalto University \And $^4$Department of CS,\\University of Helsinki} ]

%\maketitle

\input{abstract.tex}
\input{intro.tex}
\input{bn.tex} 
\input{qnml.tex}
\input{experiments.tex}
\input{conclusion}

%%\subsubsection*{Acknowledgements}

%%Use unnumbered third level headings for the acknowledgements title.
%%All acknowledgements go at the end of the paper.

%%\subsubsection*{References}

\bibliographystyle{apalike}
\bibliography{cosco}

\end{document}