### 2018-02-19 supp_mat_D.tex

\section{Prediction and Parsimony}

We took 20 UCI data sets and created 1000 permutations of each data set.
We then took the first $x$\% of each permuted data set ($x$ in
${10,20,30,\ldots,90}$) as training data and used the remaining part as
test data. We used exact structure learning to learn the best
scoring model using BDeu ($\alpha=1$), BIC, fNML and qNML,
recording the number of parameters in the learnt
models and the average predictive log-probability of each test
vector. These numbers are collected for 1000 different data
permutations, the average and variance of which appear on the following
pages, one page per data set.